Combining formal and distributional models of predicate semantics

### 2016-03-04 papers/acl2016/paper.tex
Moved the papers a bit
Former-commit-id: 52737d10d068126b1d127ac960df72f8ac9e9a99
A little more work on the paper
Former-commit-id: 2d8759dcdc0d4acf4137449484ca583a350039d4

To demonstrate these models, we focus on the task of fill-in-the-blank natural
language queries, such as ``Italian architect \blank{}''.  Similar to many
current semantic parsers, we first convert this natural language query into a
logical form: $\lambda(\entity{x}) \predicate{architect}(\entity{x}) \land
\predicate{architect\_N/N}(\entity{Italian}, \entity{x})$.  A typical semantic
parser (e.g., that of Kwiatkowski et
al.~\shortcite{kwiatkowski-2013-ontology-matching}) would then try to map this
logical form to a Freebase query, matching \entity{Italian} to
(\predicate{/people/person/nationality}, \entity{Italy}) and
\predicate{architect} to (\predicate{/type/object/type},
\entity{/architecture/architect}).  In contrast, we follow Krishnamurthy and
Mitchell~\shortcite{krishnamurthy-2015-semparse-open-vocabulary} in learning a
model for each predicate that can give a \emph{score} to potential arguments.
However, while the models of Krishnamurthy and Mitchell are purely
distributional, using Freebase only as a means of finding entities in text, we
inject additional information from Freebase into the models for each predicate.
The model we learn for \predicate{architect} has a \emph{feature} capturing the
fact that an entity has the type \entity{/architecture/architect} in Freebase,
and the model for \predicate{architect\_N/N} has a \emph{feature} capturing
whether the first argument is the country of nationality of the second
argument.

@@

@@

@@

@@

>>>>

To demonstrate these models, we focus on the task of fill-in-the-blank natural
language queries, such as ``Italian architect \blank{}''.  Similar to many
current semantic parsers, we first convert this natural language query into a
logical form: $\lambda x. \predicate{architect}(x) \land
\predicate{architect\_N/N}(\entity{Italian}, x)$.  A typical semantic
parser (e.g., that of Kwiatkowski et
al.~\shortcite{kwiatkowski-2013-ontology-matching}) would then try to map this
logical form to a Freebase query, matching \entity{Italian} to
(\predicate{/people/person/nationality}, \entity{Italy}) and
\predicate{architect} to (\predicate{/type/object/type},
\entity{/architecture/architect}).  In contrast, we follow Krishnamurthy and
Mitchell~\shortcite{krishnamurthy-2015-semparse-open-vocabulary} in learning a
model for each predicate that can give a \emph{score} to potential arguments.
However, while the models of Krishnamurthy and Mitchell are purely
distributional, using Freebase only as a means of finding entities in text, we
inject additional information from Freebase into the models for each predicate.
The model we learn for \predicate{architect} has a \emph{feature} capturing the
fact that an entity has the type \entity{/architecture/architect} in Freebase,
and the model for \predicate{architect\_N/N} has a \emph{feature} capturing
whether the first argument is the country of nationality of the second
argument.

@@

\mattnote{Add a figure here to demonstrate the logical forms.}

@@

In this section we present our models of predicate semantics.  In
\secref{formal-and-distributional} we describe how we modify Krishnamurthy and
Mitchell's (K&M's) semantic parser to include features from a formal knowledge base.
In \secref{computing-pmi} we describe how we select those features for each
predicate.  The remaining two sections then describe other improvements we made
to K&M's semantic parsing model: lexicalizing preposition and noun-compound
predicates, and using the Freebase graph to select candidate entities.

@@

The modification we make to K&M's semantic parser is simple.  For each entity
and entity pair in the data we compute an SFE feature vector $\psi(e)$ and
$\psi(e_1, e_2)$.  For each category and relation, we select a subset of the
features to have associated weights $\omega_c$ and $\omega_r$.  We then add the
weights $\omega_c$ and $\omega_r$ as parameters to be learned for each
predicate, and we add the feature vectors $\psi(e)$ and $\psi(e_1, e_2)$ as
observed features for each entity and entity pair.  The augmented category and
relation probabilities are as follows:

\begin{align*}
  P(c(e) & = \true{}) & = & \sigma ( \theta_c^T \phi_e + \omega_c^T \psi_c(e)) \\
  P(r(e_1, e_2) & = \true{}) & = & \sigma ( \theta_r^T \phi_{(e_1, e_2)} + \omega_r^T \psi_r(e_1, e_2) )
\end{align*}
where we have added subscripts on the $\psi$ functions to indicate that each
predicate selects its own set of features from the SFE feature vector $\psi$.

Note that now there are three sets of parameters to be learned by the semantic
parser: (1) $\theta$, low-dimensional distributional vectors trained for each
predicate; (2) $\phi$, low-dimensional distributional vectors trained for each
entity and entity pair; and (3) $\omega$, weights associated with the selected
formal SFE features for each predicate.  All of these parameters are optimized
jointly, using the same ranking objective used by K&M.

@@

The feature vectors returned by SFE have dimensionality in the tens of
millions.  We do not have near enough data to train that many parameters for
each of our tens of thousands of predicates, nor do we have enough RAM in our
machines to keep that many parameters in memory.  It is thus necessary to
select some small number of informative features for each predicate.  We do
this using pointwise mutual information (PMI).

PMI measures the strength of association between two variables $x$ and $y$; it
is computed as $\log\frac{p(x,y)}{p(x)p(y)}$.  In our case, we want to
associate predicates with SFE features.  However, our SFE features are for each
\emph{entity} and \emph{entity pair}, not each \emph{predicate}.  To associate
SFE features with predicates, we use the training
